---
title: "GenomicPalindromesR: mouse genome palindrome discovery with dot-bracket strings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GenomicPalindromesR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(GenomicPalindromesR)
library(dplyr)
```

## Concept

`GenomicPalindromesR` treats a DNA inverted repeat like an RNA hairpin. For a
candidate window with two complementary arms separated by a spacer, matched
positions are encoded as parentheses and unpaired positions as dots. That gives
a compact structural string such as:

```r
"((((....))))"
```

You can then search structure directly with tidyverse pipelines.

## Single-sequence scan

```{r}
hits <- find_palindromes_seq(
  sequence = "ATGCAAGCAT",
  min_arm = 4,
  max_arm = 4,
  min_loop = 2,
  max_loop = 2,
  max_mismatches = 0
)

hits
```

## Structure-first filtering

```{r}
hits %>%
  search_structure("^\\({4,}\\.{2}\\){4,}$")
```

## Mouse genome example

```{r eval = FALSE}
library(BSgenome.Mmusculus.UCSC.mm39)

mm39_hits <- find_palindromes_genome(
  genome = BSgenome.Mmusculus.UCSC.mm39,
  chromosomes = c("chr1", "chr7", "chrX"),
  min_arm = 6,
  max_arm = 12,
  min_loop = 0,
  max_loop = 20,
  max_mismatches = 1
)

mm39_hits %>%
  search_structure("^\\({6,}\\.{0,20}\\){6,}$") %>%
  summarise_palindromes()
```

## Interpretation

- more `(` and `)` means stronger inverted-repeat support
- more `.` in the arms means more broken complementarity
- larger loops can capture more hairpin-like spacer regions
